---
title: "hw03-gapminder"
author: "kioke"
date: "September 29, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tidyverse))
gapminder %>% 
  arrange(country)
```

#Tasks
Pick three of the tasks and present a table and a figure for each.

###How is life expectancy changing over time on different continents?

The figure below shows life expectancy over time, separated by continent by the colour of the points.
```{r}
gapminder %>% 
  ggplot(aes(year,lifeExp,colour=continent)) +
  geom_jitter(alpha=0.33) +
  labs(x = "Year", y = "Life Expectancy") #To change the axis labels

```
We could also choose to represent the different continents on different plots by facetting the data, if the above figure is too visually busy to read.
```{r}
gapminder %>% 
  ggplot(aes(year,lifeExp)) +
  geom_jitter(alpha=0.2) +
  facet_wrap( ~ continent) +
  labs(x = "Year", y = "Life Expectancy")
```
Below is a table summarizing the above figures globally.
```{r}
gapminder %>% 
  group_by(year) %>% 
  summarize(mean_life_exp = mean(lifeExp))

```
We can also look at the means over time in a table with continents as another variable.
```{r}
gapminder %>% 
  group_by(year,continent) %>% 
  summarize(mean_life_exp = mean(lifeExp))

```


###Mean Life Expectancy, 1982-1997
In order to compute the mean of any data between 1982 and 1997, we first need to filter out other years
```{r}
mill <- gapminder %>% 
  filter(year >= 1982 & year <= 1997)
```
Now to determine the mean life expectancy of those years.
```{r}
mill %>% 
  group_by(continent) %>% 
  summarize(mean_mill_life_exp = mean(lifeExp))
```


```{r}
mill %>% 
  ggplot(aes(gdpPercap, lifeExp)) +
  geom_point() + 
  scale_x_log10()
```
```{r}
mill %>% 
  group_by(continent) %>% 
  summarize(mean_mill_life_exp = mean(lifeExp)) %>% 
  ggplot(aes(continent, mean_mill_life_exp)) +
  geom_point(size = 3) +
  labs(x = "Continent", y = "Mean Life Expectancy 1982-1997")
```


