---
title: "hw03-gapminder"
author: "kioke"
date: "September 29, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tidyverse))
```

#Tasks
Pick three of the tasks and present a table and a figure for each.

###How is life expectancy changing over time on different continents?

The figure below shows life expectancy over time, separated by continent by the colour of the points.
```{r}
gapminder %>% 
  ggplot(aes(year,lifeExp,colour=continent)) +
  geom_jitter(alpha=0.33) +
  labs(x = "Year", y = "Life Expectancy") #To change the axis labels

```
We could also choose to represent the different continents on different plots by facetting the data, if the above figure is too visually busy to read.
```{r}
gapminder %>% 
  ggplot(aes(year,lifeExp)) +
  geom_jitter(alpha=0.2) +
  facet_wrap( ~ continent) +
  labs(x = "Year", y = "Life Expectancy")
```

```{r}

  
```

```{r}
library(gapminder)
library(tidyverse)

```

```{r}
dim(gapminder)
str(gapminder)
```

```

