---
title: "hw03-gapminder"
author: "kioke"
date: "September 29, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tidyverse))
```

#Tasks
I explored the Gapminder data set to manipulate and visualize the data contained within. Specifically, I examined life expectancy over time for different continents, the mean life expectancy in a given time period, and looked at how historical events might be represented in this dataset.

###How is life expectancy changing over time on different continents?

The figure below shows life expectancy over time, separated by continent by the colour of the points.
```{r}
gapminder %>% 
  ggplot(aes(year,lifeExp,colour=continent)) +
  geom_jitter(alpha=0.33) +
  labs(x = "Year", y = "Life Expectancy") #To change the axis labels

```


We could also choose to represent the different continents on different plots by facetting the data, if the above figure is too visually busy to read. The figure below separates the data by continent into different plots.

```{r}
gapminder %>% 
  ggplot(aes(year,lifeExp)) +
  geom_jitter(alpha=0.2) +
  facet_wrap( ~ continent) +
  labs(x = "Year", y = "Life Expectancy")
```


Below is a table summarizing the above figures globally.

```{r}
gapminder %>% 
  group_by(year) %>% 
  summarize(mean_life_exp = mean(lifeExp)) %>% 
  knitr::kable()

```

We can also look at how the means change over time in a table with continents as another variable.

```{r}
gapminder %>% 
  group_by(year,continent) %>% 
  summarize(mean_life_exp = mean(lifeExp)) %>% 
  knitr::kable()

```


###Mean Life Expectancy, 1982-1997

In order to compute the mean of any data between 1982 and 1997, we first need to filter out other years

```{r}
mill <- gapminder %>% 
  filter(year >= 1982 & year <= 1997)
```

Now to determine the mean life expectancy of those years.

```{r}
mill %>% 
  group_by(continent) %>% 
  summarize(mean_mill_life_exp = mean(lifeExp)) %>% 
  knitr::kable()
```

We could also plot the data in the above table.

```{r}
mill %>% 
  group_by(continent) %>% 
  summarize(mean_mill_life_exp = mean(lifeExp)) %>% 
  ggplot(aes(continent, mean_mill_life_exp)) +
  geom_point(size = 3) +
  labs(x = "Continent", y = "Mean Life Expectancy 1982-1997")

```


###The Soviet-Afghan War
In late 1979, the Soviet Union invaded Afghanistan and kept troops deployed there until the late 1980s. Perhaps the effects of this invasion and occupation will be apparent in the Gapminder dataset.

```{r}
gapminder %>% 
  filter(country == "Afghanistan") %>% 
  mutate(popmil = round(pop/10^6,2)) %>% 
  ggplot(aes(size=gdpPercap, year, popmil)) +
  geom_point() +
  labs(x = "Year", y = "Population (millions)")
```

The data shows a noticeable drop in population in the timeframe of the war in Afghanistan, though effects on GDP per capita and life expectancy were not apparent in the given data. How does the drop in population in Afghanistan between 1977 and 1982 compare to periods of population decline in other countries in Asia?


```{r}
gapminder %>%
  mutate(changepop = pop - lag(pop)) %>%
  mutate(percent_change = round(changepop/pop*-100,2)) %>% 
  arrange(changepop) %>% 
  filter(changepop < 0 & continent == "Asia", year!=1952) %>% #Some countries had negative change values for the year 1952 because the lag() function subtracted their 1952 population from the 2007 population of another country
  select(-gdpPercap, -lifeExp) %>% 
  knitr::kable()
```

